Towards a miniKanren with fair search strategies

### 2019-05-24 Paper/paper.tex
cite Dmitris

@@

\NOTE[LKC]{We actually have two new BFS implementations, an easy one and an 
improved one (BFSimp/BFSopt). Both of them runs faster than the current BFS. I 
didn't name the easy one because I considered it just a stepstone to BFSimp. 
But now I think giving names to both of them would be more clear.}

@@

miniKanren is a family of relational programming languages.
\citet{Friedman:2005:RS:1121583,friedman_reasoned_2018} introduce
miniKanren and its implementation in \emph{The Reasoned Schemer} 
and \emph{The Reasoned Schemer, 2nd Ed} (TRS2). \citet{byrd2017unified} have 
demonstrated that miniKanren programs are useful in solving several difficult
problems. miniKanren.org contains the seeds of many difficult problems and 
their solutions.

\NOTE[DPF]{...  Mostly the point is to also include a reference to
the two dimitris and to the first OCamren (OCanren), whatever. It is up to you
in this paragraph.  ...}
\NOTE[LKC]{I have fixed the cite to TRS and TRS2.}
\NOTE[LKC]{What is the point of citing dmitris?}
\NOTE[LKC]{Are we citing OCanren to show that miniKanren is widely used? If 
yes, do we need to cite other implementations as well?}

\NOTE[DPF]{I have decided to agree with you that we should not use 
BFSser, so I am going to remove all occurrences.  
Also, we should say, ``Our implementation of BFS''
and then say, shortened to ``Our BFS'' and then you can use it in the table,
so one column should be Our BFS and the other should be BFS.}

\NOTE[DPF]{Occurrences of BFS (without Our) mean Seres et al. (We need only say
that once, or very occasionally}

@@

To summarize our contributions, we
\begin{itemize}
	\item propose and implement \textbf{b}alanced \textbf{i}nterleaving 
depth-first search (\DFSbi{}), a new search strategy with almost-fair \disj{}.
	\item propose and implement \textbf{f}air depth-first search (\DFSf{}), 
a new search strategy with fair \disj{}.
	\item implement in a new way the standard breath-first search (\BFS), a 
search strategy with fair \disj{} and fair \conj{}. We refer to ``the current 
BFS implementation'' (``BFS'' for short) as the implementation by 
\citet{seres1999algebra} and refer to ``our BFS implementation'' (``our BFS'' 
for short) as our new one. We formally prove that the two BFS implementations 
are semantically equivalent, however, our BFS runs faster in all benchmarks and 
is shorter.
\end{itemize}

@@

In this section, we define fair \disj{}, almost-fair \disj{} and fair \conj{}. 
Before going further into fairness, we give a short review of the terms:
\emph{state}, search \emph{space}, and \emph{goal}.
A \emph{state} is a collection of constraints. (Here, we restrict 
constraints to unification constraints.) Every answer corresponds to a 
state. A space is a collection of states. And a \emph{goal} is a 
function from a state to a space.

@@

\NOTE[LKC]{I have pointed out ``\DFSf{} also has the space extravagance'' at 
the end of the above paragraph.}

@@

\NOTE[LKC]{I change the benchmark '(I love you) such that it doesn't prevent 
closures from getting into pairs because I consider this restriction not 
reasonable. As for quines, they are already very slow so I didn't make the 
changes.}

@@

\NOTE[LKC]{I rewrote the following paragraph and the table to include the 
additional BFS and update the '(I love you) benchmark.

Also, all running time are re-measured by running in command-line. I used to 
run benchmarks with DrRacket. Thus, the time measurements are smaller.}

In general, only \DFSi{} and \DFSbi{} constantly perform well. \DFSf{}
is just as efficient in \appendo, \reverso, and \texttt{quine}s. All BFS
implementations have obvious overhead in most cases. 
Among the three variants of DFS, which all have unfair \conj{}, \DFSf{} is most 
resistant to clause permutation in \texttt{quines}, followed by \DFSbi{} then 
\DFSi{}. \DFSf{}, however, runs out of memory in \texttt{'(I love you)}s. In 
contrast, \DFSbi{} still performs more stable than \DFSi{}. Thus, we consider 
\DFSbi{} a competitive alternative to \DFSi{}, the current standard miniKanren 
search strategy. 
Among the three implementations of \BFS, our \BFS{}s constantly perform as well 
or better. \BFSimp{} only out-performs \BFSuni{} significantly in 
\veryrecursiveo{} 300000, while performing as well or worse elsewhere. The 
reason might be that the size of bags are usually too small to be worried about.
It is hard to conclude how \disj{} fairness affect performance based on these 
statistics. Fair \conj{}, however, imposes considerable overhead constantly 
except in \appendo. The reason might be that those strategies tend to keep more 
intermediate answers in the memory.

@@

\NOTE[DPF]{We might find the reference Ocanren in the two Dmitri's paper
and say something here about it.}

@@

We devise two new search strategies (i.e. balanced interleaving DFS 
(\DFSbi{}) and fair DFS (\DFSf{})) and devise a new 
implementation of \BFS. These strategies have different features 
in fairness: \DFSbi{} has an almost-fair \disj{} and unfair \conj{}. 
\DFSf{} has fair \disj{} and unfair \conj{}. \BFS{} has both fair
\disj{} and fair \conj{}.

Our quantitative evaluation shows that \DFSbi{} and \DFSf{} are competitive 
alternatives to \DFSi{}, the current miniKanren search strategy,
and that \BFS{} is less practical.
% and that \BFSser{} is less practical than other breadth-first strategies.

>>>>

\NOTE[LKC]{
We actually have two new BFS implementations. One is modified from DFSf with 
two simple steps. It has been described briefly at the beginning of the BFS 
section but was not given a name. The other is the one described with two large 
figures and was given the name ``our improved BFS''. Both of them runs faster 
than the current BFS.

I didn't name the first one because (1) it is so simple and (2) I considered 
it merely a stepstone to our improved BFS. Now I decided to also give the first 
BFS implementation a name. This is because I think having a nameless impl is 
confusing.

Currently, the nameless implementation is given the name ``our BFS''. And I 
have changed the paper accordingly.

BUT, I am thinking about going further --- removing ``our improved BFS'' from 
this paper. The first reason is that our BFS is actually more efficient than 
our improved BFS in most benchmarks (see Table 1), despite its theoretical 
space extravagence. The second reason is that DFSi, DFSf, and our BFS look nice 
when sitting together.}

@@

@@

miniKanren is a family of relational programming languages.
\citet{Friedman:2005:RS:1121583,friedman_reasoned_2018} introduce
miniKanren and its implementation in \emph{The Reasoned Schemer} 
and \emph{The Reasoned Schemer, 2nd Ed} (TRS2). Although originally written in 
Scheme, miniKanren has been ported to many other languages (e.g. 
OCanren\citep{kosarev2018typed}). \citet{byrd2017unified} have 
demonstrated that miniKanren programs are useful in solving several difficult
problems. miniKanren.org contains the seeds of many difficult problems and 
their solutions.

\NOTE[DPF]{...  Mostly the point is to also include a reference to
the two dimitris and to the first OCamren (OCanren), whatever. It is up to you
in this paragraph.  ...}
\NOTE[LKC]{I have fixed the references to TRS and TRS2.}
\NOTE[LKC]{I have cited OCanren in this paragraph.}

\NOTE[DPF]{I have decided to agree with you that we should not use 
BFSser, so I am going to remove all occurrences.  
Also, we should say, ``Our implementation of BFS''
and then say, shortened to ``Our BFS'' and then you can use it in the table,
so one column should be Our BFS and the other should be BFS.}
\NOTE[DPF]{Occurrences of BFS (without Our) mean Seres et al. (We need only say
that once, or very occasionally}

@@

To summarize our contributions, we
\begin{itemize}
	\item propose and implement \textbf{b}alanced \textbf{i}nterleaving 
depth-first search (\DFSbi{}), a new search strategy with almost-fair \disj{}.
	\item propose and implement \textbf{f}air depth-first search (\DFSf{}), 
a new search strategy with fair \disj{}.
	\item implement in two new ways the standard breath-first search (\BFS), a 
search strategy with fair \disj{} and fair \conj{}. We refer to ``the current 
BFS implementation'' (``BFS'' for short) as the implementation by 
\citet{seres1999algebra} and refer to ``our BFS implementation'' (``our BFS'' 
for short) and ``our improved BFS implementation'' (``our improved BFS`` for 
short) as our new ones. We also refer to ``our \BFS{}s'' to both \BFSuni{} and 
\BFSimp{}. We argue that our \BFS{}s are semantically equivalent and formally 
prove that \BFSimp{} and \BFSser{} are semantically equivalent, however, our 
\BFS{}s run faster in all benchmarks and are shorter.
\end{itemize}

@@

In this section, we define fair \disj{}, almost-fair \disj{} and fair \conj{}. 
Before going further into fairness, we give a short review of the terms:
\emph{state}, search \emph{space}, and \emph{goal}.
A \emph{state} is a collection of constraints. (Here, we restrict 
constraints to unification constraints.) Every answer corresponds to a 
state. A \emph{space} is a collection of states. And a \emph{goal} is a 
function from a state to a space.

@@

@@

\NOTE[LKC]{I have changed the benchmarks about '(I love you), such that their 
relational interpreters doesn't prevent closures from getting into pairs. I did 
so because I consider this restriction doesn't make sense. Benchmarks about 
quine are already very slow so I didn't apply similar changes.}

@@

\NOTE[LKC]{I rewrote the following paragraph and the table to include our BFS 
in additional to our improved BFS.}
\NOTE[LKC]{I also re-measured all running time. I used to run benchmarks with 
DrRacket, now I do it in command-line. Thus, the time measurements are smaller 
and more stable.}

In general, only \DFSi{} and \DFSbi{} constantly perform well. \DFSf{}
is just as efficient in \appendo, \reverso, and \texttt{quine}s. All BFS
implementations have obvious overhead in most cases. 
Among the three variants of DFS, which all have unfair \conj{}, \DFSf{} is most 
resistant to clause permutation in \texttt{quines}, followed by \DFSbi{} then 
\DFSi{}. \DFSf{}, however, runs out of memory in \texttt{'(I love you)}s. In 
contrast, \DFSbi{} still performs more stable than \DFSi{}. Thus, we consider 
\DFSbi{} a competitive alternative to \DFSi{}, the current standard miniKanren 
search strategy. 
Among the three implementations of \BFS, our \BFS{}s constantly perform as well 
or better. Among our \BFS{}s, \BFSimp{} only performs better in 
\veryrecursiveo{} 300000. The reason might be that the size of bags are usually 
too small to be worried about. 
It is hard to conclude how \disj{} fairness affect performance based on these 
statistics. 
Fair \conj{}, however, imposes considerable overhead constantly except in 
\appendo. The reason might be that those strategies tend to keep more 
intermediate answers in the memory.

@@

\citet{rozplokhas2018improving} address the non-commutativity of conjunction, 
while our work about \disj{} fairness addresses the non-commutativity of 
disjunction.

\NOTE[DPF]{We might find the reference Ocanren in the two Dmitri's paper
and say something here about it.}
\NOTE[LKC]{added}

@@

We devise two new search strategies (i.e. balanced interleaving DFS 
(\DFSbi{}) and fair DFS (\DFSf{})) and devise two new 
implementations of \BFS. These strategies have different features 
in fairness: \DFSbi{} has an almost-fair \disj{} and unfair \conj{}. 
\DFSf{} has fair \disj{} and unfair \conj{}. \BFS{} has both fair
\disj{} and fair \conj{}.

Our quantitative evaluation shows that \DFSbi{} is a competitive 
alternative to \DFSi{}, the current miniKanren search strategy,
and that \BFS{} is less practical.
% and that \BFSser{} is less practical than other breadth-first strategies.
